Unrestricted Termination and Non-Termination Arguments for Bit-Vector Programs

### 2015-04-12 slides/esop-termination.tex
Tautologies

This formula is satisfiable iff there exists a recurrence set $\blue{S}$ containing
an initial state $\red{x_0}$, with successor function $\green{N}$, i.e. iff the loop
does not terminate for some input $\red{x_0}$.
\end{frame}

@@

>>>>

This formula is satisfiable iff there exists a recurrence set $\blue{S}$ containing
an initial state $\red{x_0}$, with successor function $\blue{N}$, i.e. iff the loop
does not terminate for some input $\red{x_0}$.
\end{frame}

@@

\begin{frame}{The Power of Tautologies}
 Our second-order solver works by trying to synthesise programs for each of the existentially quantified functions.

 This problem is NEXPTIME-hard, but in practice is quite fast if there is some short program that satisfies
 the constraints.

 However if the constraints are UNSAT, the solver will take a very long time to discover this.
\end{frame}

\begin{frame}{The Power of Tautologies}
 For a given loop, the formula \red{[CT]} is satisfiable iff the loop terminates and the formula \blue{[SNT]} is
 satisfiable iff it does not terminate.  Therefore, the formula
 \[
  \red{\text[CT]} \vee \blue{\text[SNT]}
 \]
 is satisfiable for \emph{every} loop.

 Since our solver produces witnesses to show that a formula is satisfiable, it will terminate with enough
 information to check whether the \red{[CT]} or \blue{[SNT]} disjunct has been satisfied.

 This allows us to avoid the expensive case of having to check an UNSAT formula.
\end{frame}
